---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

## Libraries
```{r Setup, echo=TRUE, include=FALSE}
library("lubridate")
library("dplyr")
library("tidyr")
library("ggplot2")
```

## Loading and preprocessing the data
```{r Load, cache=TRUE}
setwd("~/COURSERA/DATA SCIENTIST CERTIFICATE/COURSE 5/WEEK 1/R")
activity <- read.csv("activity.csv")
activity$date <- as_date(activity$date)
```

## What is mean total number of steps taken per day?
```{r meanSteps}
x<-group_by(drop_na(activity), date)%>%summarise(sum = sum(steps, na.rm = TRUE))
hist(x$sum, breaks = 10)
```
The mean equals `r mean(x$sum)` and the median steps per day is `r median(x$sum)`

## What is the average daily activity pattern?
```{r averageDaily}
w<-group_by(drop_na(activity), interval)%>%summarise(mean = mean(steps))
plot(w$interval, w$mean, type = "l", xlab = "interval", ylab = "average #steps")
```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r whichInterval}
w$interval[which.max(w$mean)]
```

## Imputing missing values
```{r imputting}
#Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
sum(is.na(activity$steps))

# Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

# activity$steps[is.na(activity$steps)]<-y Assign mean to all NAs

# Create a new dataset that is equal to the original dataset but with the missing data filled in.
w <- activity
w2<-group_by(w, interval)%>%summarise(mean = mean(steps,na.rm = TRUE))

left_join(w, w2, by="interval") ->m

m$steps[is.na(w$steps)]<-m$mean[is.na(w$steps)]

group_by(m, date)%>%summarise(sum = sum(steps))->n

hist(n$sum, breaks = 10)

```

## Are there differences in activity patterns between weekdays and weekends?
```{r dayDifference}
v <- mutate(activity, daysofweek = weekdays(activity$date, abbreviate = TRUE))

# Create factor variable with two levels: weekday and weekend.
grepl("S(at|un)", v$daysofweek)->v$isweekend

# Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

v2<-group_by(v, interval, isweekend)%>%summarise(mean = mean(steps, na.rm = TRUE))

par(mfrow=c(1,2))
plot(v2$interval[v2$isweekend], v2$mean[v2$isweekend], type = "l", ylab = "Steps", xlab = "5 min Intervals on Weekend")

plot(v2$interval[!v2$isweekend], v2$mean[!v2$isweekend], type = "l", ylab = "Steps", xlab = "5 min Intervals on Weekday")
```